\myTitle

### 2024-05-22 body-text-of-paper.tex
Eric completed the edits from Tony's comments

Genetic clustering algorithms are commonly used to describe genetic structure among populations, reflecting the cumulative
effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations
is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker
set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently
interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location
and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications,
such as influencing evolutionary processes, maintaining genetic variation through gene flow, driving disease dynamics and affecting patterns
of invasive species expansion \citep{huestis2019windborne, estoup2010reconstructing}.

@@

A best practice for evaluating the behavior of genetic clustering algorithms on empirical data is to conduct
parallel analyses on simulated data that mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine} while providing
the benefit of known parameters (e.g.~$F_\mathrm{ST}$, admixture fractions, etc.) that can
be compared to estimates.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}).
However, sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
samples, and
inflates the perceived statistical power to correctly assign individuals (or segments of the genome) back to source populations.
We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
\citet{anderson2008improved} used simulations to document this type of artifact in the context of
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.

@@

In this case,
each set of $N$ individuals is a sample from exactly the same population, so
there is clearly no basis for performing population assignment between $A$ and $B$;
however, we treated samples $A$ and $B$ as if they were sampled from two potentially different
populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
$q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$.
We simulated these genotypes according to the "admixture model" used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each
locus, independently, the origin
of each gene copy was simulated to be from
population $A$ with probability $q_A$ and from population $B$ with probability $1-q_A$.  Subsequently,
the allelic types of gene copies from $A$ ($B$), at a locus, were simulated by sampling with replacement
from the alleles at that locus in sample $A$ ($B$).
We then analyzed all the genotypes ($9n + N + N$) using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original
$N$ samples from each
population were provided as learning samples, and also an
unsupervised analysis when the origin of the samples from $A$ and $B$ were regarded as unknown.

@@

We also simulated the genomes of 48,000 individuals of each
hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
below, with 18 chromosomes (2.35 gigabases of physical genome length)  and
sex-averaged recombination rates
for pigs, as estimated by from \citet{tortereau2012high}. In this case, we recorded the
admixture fraction of each simulated individual as the cumulative length of all genome segments
that originated from species~$a$, divided by the total genome length.  This provides the distribution of the true admixture fraction
when physical linkage of markers is not ignored.

@@

\citeauthor{rizza2023limited} mapped the 65 markers to the Omyk\_v1.0 genome assembly (RefSeq 
Accession: GCF\_002163495.1), so we used
the linkage map described in \citep{pearse2019sex} that was prepared by linkage mapping 
of $\approx$57,000 markers described in \citet{palti2015development} to Omyk\_v1.0. We mapped these $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
of 32,923 markers throughout the genome and rendered it to plink format for use with \gscramble{}.

@@

We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
provided empirical examples with trout and wild pig data, we demonstrate the basics of 
creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
 to preserve founder individuals' genotypes in contrast to \textit{a priori} shuffling alleles 
within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
is important to note that each time a gamete is segregated in a GSP, its complement is also 
segregated---an example of antithetical variate sampling in Monte Carlo methods \citep{hammersley}.
Thus, in the  example of relatedness, fixed specified founders should be repeatedly "mated" with one offspring 
sampled per iteration to avoid sampling of anticorrelated genotypes.

>>>>

Genetic clustering algorithms are commonly used to describe genetic structure among populations, reflecting the cumulative
effects of isolation, connectivity (i.e., genetic migration), selection, and genetic drift.  The ability of such algorithms to differentiate populations
is limited by the extent of genetic differentiation among populations and the capacity to resolve those differences given the available marker
set and sample sizes from the respective populations.  Based on an initial description of genetic structure, ecologists are frequently
interested in describing patterns of genetic movement among populations, often inferred from the mismatch between the sampling location
and genetic origins of a given individual \citep{paetkau1995microsatellite,wilson2003bayesian}.  Genetic movement has many implications,
such as influencing evolutionary processes, maintaining genetic variation through gene flow, driving disease dynamics, and affecting patterns
of invasive species expansion \citep{huestis2019windborne, estoup2010reconstructing}.

@@

A best practice for evaluating the behavior of genetic clustering algorithms on empirical data is to conduct
parallel analyses on simulated data that mirror the empirical data set
as closely as possible \citep{vaha2006efficiency,anderson2008improved,latch2011fine} while providing
the benefit of known parameters (e.g.~$F_\mathrm{ST}$, admixture fractions, etc.) that can
be compared to estimates.
Techniques to simulate individual genotypes in this context frequently use sampling with replacement from the
observed allele frequencies (see, for example, \citealt{nielsen2006hybridlab,kinziger2008hybridization}). In this
approach, allele frequencies are estimated for each population (or cluster) from the samples, and then new
genotypes are simulated at each locus by drawing, for each simulated diploid individual, two gene copies from a 
multinomial distribution with two trials
and cell probabilities given by the estimated allele frequencies.  Such a simulation is akin to assuming that the
population is of infinite size with allele frequencies given by those observed in the sample; however the allele
frequencies in the sample differ from those in the actual population because of the variation from random sampling.
As a consequence, this sort of sampling with replacement implicitly increases the pairwise genetic distance among simulated samples, relative to the observed
samples, and
inflates the perceived statistical power to correctly assign individuals (or segments of the genome) back to source 
populations.  In effect, simulations that sample with replacement from the samples are assessing how well one could 
assign individuals to populations
that have the same allele frequencies as the samples, rather than the same allele frequencies as the populations
{\em from which the samples were drawn}, which is what is needed.
We refer to this phenomenon as resampling-induced, spurious power inflation (RISPI).
\citet{anderson2008improved} used simulations to document this type of artifact in the context of
genetic stock identification (GSI), a type of clustering application in which individuals are constrained to have
all their ancestry from a single subpopulation.

@@

In this case,
each set of $N$ individuals is a sample from exactly the same population, so
there is clearly no basis for performing population assignment between $A$ and $B$;
however, we treated samples $A$ and $B$ as if they were sampled from two potentially different
populations, and simulated $9n$ new individuals: $n$ new individuals at each of the 9 values of
$q_A$, the admixture fraction for population $A$, in  $\{0, \frac{1}{8}, \frac{1}{4}, \frac{3}{8}, \frac{1}{2}, \frac{5}{8}, \frac{3}{4}, \frac{7}{8}, 1\}$.
We simulated these genotypes according to the ``admixture model'' used in {\em structure} \citep{pritchard2000inference}.  Briefly, at each
locus, independently, the origin
of each gene copy was simulated to be from
population $A$ with probability $q_A$ and from population $B$ with probability $1-q_A$.  Subsequently,
the allelic types of gene copies from $A$ ($B$), at a locus, were simulated by sampling with replacement
from the alleles at that locus in sample $A$ ($B$).
We then analyzed all the genotypes ($9n + N + N$) using ADMIXTURE with $K=2$. On each data set we performed a supervised analysis in which the original
$N$ samples from each
population were provided as learning samples, and also an
unsupervised analysis when the origin of the samples from $A$ and $B$ were regarded as unknown.

@@

We also simulated the genomes of 48,000 individuals of each
hybrid category using \gscramble{}'s {\footnotesize\tt segregate()} function, as described
below, taking as an example, a species with 18 chromosomes across 2.35 gigabases of total physical genome length
and sex-averaged recombination rates the same as for pigs, as estimated by from \citet{tortereau2012high}. (We chose
a pig-like genome here, as wild pigs form one of our example empirical data sets for illustration, below.)  In this case, we recorded the
admixture fraction of each simulated individual as the cumulative length of all genome segments
that originated from species~$a$, divided by the total genome length.
This provides the distribution of the true admixture fraction
when physical linkage of markers is not ignored.

@@

\citeauthor{rizza2023limited} mapped the 65 markers to the Omyk\_v1.0 genome assembly (RefSeq 
Accession: GCF\_002163495.1), so we used
the linkage map described in \citep{pearse2019sex} that was prepared 
with $\approx$57,000 markers described in \citet{palti2015development} mapped to Omyk\_v1.0. We mapped these $\approx$57,000 markers to Omyk\_v1.0, and then resolved inconsistencies to produce a map
of 32,923 markers throughout the genome and rendered it to plink format for use with \gscramble{}.

@@

We envision \gscramble{} to be relevant for a diverse range of potential applications. In our 
provided empirical examples with trout and wild pig data, we demonstrate the basics of 
creating hybrid classes. In much the same manner, \gscramble{} has been used to evaluate the statistical power 
of hybrid classification used to differentiate wild boar hybrids from domestic pigs 
\citep{smyser2024probgen}. Specifically \citet{smyser2024probgen} used the segregation options in \gscramble{} 
 to preserve founder individuals' genotypes in contrast to \textit{a priori} shuffling alleles 
within populations. This retains the subpopulation structure present within certain populations---thereby, better reflecting the 
actual admixture variation expected from wild boar and domestic pig hybrids. We think \gscramble{} 
may also be useful for assessing relatedness estimates using pedigrees of varying complexity. Here, it 
is important to note that each time a gamete is segregated in a GSP, its complement is also 
segregated---an example of antithetical variate sampling in Monte Carlo methods \citep{hammersley}.
Thus, in the  example of relatedness, fixed specified founders should be repeatedly ``mated'' with one offspring 
sampled per iteration to avoid sampling of anticorrelated genotypes.
